Seminar04 - two group comparisons and data aggregation
========================================================

Read in the photoRec dataset
------

```{r}
prDat <- read.table("../photoRecDataset/dataClean/GSE4051_data.tsv")
prDes <- readRDS("../photoRecDataset/dataClean/GSE4051_design.rds")
str(prDat, max.level=0)
str(prDes)
```

Two-sample tests - one gene
-----

Let's extract the data for one gene and put in a data.frame with the experimental information.

```{r}
set.seed(987)
(theGene <- sample(1:nrow(prDat), 1))

pDat <- data.frame(prDes, genExp=unlist(prDat[theGene,]))
str(pDat)
```

### explore pDat

Always explore the data before plunging into analysis! What are the sample means in the wild type and Nrl knockout groups (yes, we're ignoring developmental stage today)? **aggregate()** and other data aggregation functions are explained below.

```{r}
aggregate(genExp ~ gType, pDat, FUN=mean)
```

Let's make a stripplot so we can sanity test our t test result.

```{r}
library(lattice)
stripplot(gType ~ genExp, pDat)
```

### We will do a two-sample t test comparing wild type to the Nrl knockouts.

```{r}
t.test(genExp ~ gType, pDat)
```

If we save the t test result, we can inspect what it is.

```{r}
ttRes <- t.test(genExp ~ gType, pDat)
str(ttRes)
```

extract test statistic and p- value

```{r}
ttRes$statistic
ttRes$p.value
```

The t- test with equal variance assumption

```{r}
ttResEqlVar <- t.test(genExp~gType, pDat, var.equal=T)
str(ttResEqlVar)

```

the Wilcoxon test for differential expression

```{r}
wtRes <- wilcox.test(genExp ~ gType, data=pDat)
str(wtRes)
```

the Kolmogorov-Smirnov test for differential expression

```{r}
kstRes  <- ks.test(pDat$genExp[pDat$gType == "wt"], pDat$genExp[pDat$gType=="NrlKO"])
str(kstRes)
```

generate a table with test statistic and p values

```{r}
tests <- list(test_stat=c(ttRes$statistic, ttResEqlVar$statistic, wtRes$statistic, kstRes$statistic), pvalue=c(ttRes$p.value, ttResEqlVar$p.value, wtRes$p.value, kstRes$p.value))

testsDF <- as.data.frame(tests)

str(testsDF)
names(testsDF)
rownames(testsDF)  <- c("t.test", "t.test_EV", "wilcox.test", "KS.test")
testsDF

```


